Binding of cations to phospholipid bilayers

### 2015-02-09 LIPIDionINTERACT/LIPIDionINTERACT.tex
update lipid-ion interaction manuscript

%\author{Patrick Fuchs}
%\author{Matti Javanainen}
%\author{Antti Lamberg}
%\author{Alexander Lyubartsev}
%\author{Markus S. Miettinen}
%\author{Luca Monticelli}
%\author{Jukka Määttä}
%\setcounter{page}{1}
\author{O. H. Samuli Ollila}
\footnote{Author to whom 
correspondence may be addressed. E-mail: samuli.ollila@aalto.fi.\\
}
\affiliation{
  Helsinki Biophysics and Biomembrane Group,
Department of Biomedical Engineering and Computational Science, Aalto University, Espoo, Finland
}
%\author{Marius Retegan}
%\author{Hubert Santuz}
%\author{Joona Tynkkynen}
%\author{Alexander Vogel}

@@

Slipid\cite{jambeck12}, ionFF~\cite{??}\todoi{Andrea Catte, please check the numbers and let us know if you share some files through Zenodo}    &   DPPC & 150 (NaCl) & ? & ? & ? & ?  & ? & 323 & ? & ? &?  \\

@@

\subsubsection{Slipids}
The simulation without ions is the same as in~\cite{botan15}.
\todo{Andrea Catte, please deliver the details.}

>>>>

\author{Andrea Catte}
\thanks{University of East Anglia, Norwich, United Kingdom}
\author{Matti Javanainen}
\thanks{Tampere University of Technology, Tampere, Finland}
\author{Markus S. Miettinen}
\thanks{Fachbereich Physik, Freie Universit\"at Berlin, Berlin, Germany}
%\setcounter{page}{1}
\author{O. H. Samuli Ollila} 
\thanks{{\bf Author to whom correspondence may be addressed. E-mail: samuli.ollila@aalto.fi.}}
\thanks{Helsinki Biophysics and Biomembrane Group, Department of Biomedical Engineering and Computational Science, Aalto University, Espoo, Finland}
%\footnote{Author to whom 
%correspondence may be addressed. E-mail: samuli.ollila@aalto.fi.\\
%}
%\affiliation{
%  Helsinki Biophysics and Biomembrane Group,
%Department of Biomedical Engineering and Computational Science, Aalto University, Espoo, Finland
%}

@@

Slipid\cite{jambeck12}, ionFF~\cite{??}\todoi{Andrea Catte, please let us know if you share some files through Zenodo}    &   DPPC & 150 (NaCl) & 600 & 18000 & 49 & 0  & 49 & 323 & 100 & 40 &?  \\

@@

\subsubsection{Slipids}
The simulation without ions is the same as in~\cite{botan15}.
\todo{DONE}

For the simulations with ions, the starting DPPC lipid bilayer, which was built with the online CHARMM-GUI
(http://www.charmm-gui.org/), contained 600 lipids, 30 water molecules/lipid, Na+ and Cl- ions (150 mM NaCl). 
The TIP3p water model was used to solvate the system. AA MD simulations of DPPC lipid bilayers were performed 
at ten different temperatures (283, 298, 303, 308, 312, 313, 314, 318, 323, and 333 K) using the GROMACS software 
package version 4.5.5 and the Stockholm lipids (Slipids) force field parameters for phospholipids. After energy 
minimization and a short equilibration run of 50 ps (time step 1fs), 100 ns production runs were performed using 
a time step of 2 fs with leap-frog integrator. All covalent bonds were constrained with the LINCS
algorithm. Coordinates were written every 100 ps. PME with real space cut-off at 1.0 nm was used for Coulomb 
interactions. Lennard-Jones (LJ) interactions were switched to zero between 1.0 nm and 1.4 nm. The neighbour 
lists were updated every 10th step with a cut-off of 1.6 nm. Temperature was coupled separately for upper and 
bottom leaflets of the lipid bilayer, and for water to one of the temperatures reported above with the Nosè-Hoover 
thermostat using a time constant of 0.5 ps. Pressure was semi-isotropically coupled to the atmospheric pressure 
with the Parrinello-Rahman barostat using a time constant of 10 ps.
The last 40 ns of each simulation was employed for the analysis of DPPC choline and glycerol backbone order parameters.
